RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2023-07-28 main.tex
:speech_balloon: Update system model

% glossaries-extra
\glsdisablehyper
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communication}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communication}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmmac}{DMMAC}{Discrete Memoryless Multiple Access Channel}
\newacronym{dcmc}{DCMC}{Discrete-input Continuous-output Memoryless Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{mbc}{MBC}{Monostatic \glsentryshort{bc}}
\newacronym{bbc}{BBC}{Bistatic \glsentryshort{bc}}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{nlos}{NLoS}{Non-Line-of-Sight}
\newacronym{zf}{ZF}{Zero-Forcing}
\newacronym{mmse}{MMSE}{Minimum Mean-Square-Error}
\newacronym{fpga}{FPGA}{Field-Programmable Gate Array}
\newacronym{ber}{BER}{Bit Error Rate}

@@

	\emph{Second,} we address the first challenge and propose a practical receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern, and coherently decodes the primary link.
	We consider backscatter energy detection and formulate a \gls{dmmac} over disjoint decision regions.
	Thanks to the double modulation and symbol-level passive precoding, the semi-coherent backscatter decoding is essentially part of channel training.
	Once the backscatter information is successfully decoded, its contribution can be modelled within composite channel as dynamic passive beamforming (rather than re-precoded and cancelled).
	It requires only one energy comparison and re-encoding per backscatter symbol and is much efficient than conventional decoding based on \gls{sic} and \gls{mrc}.
	It is also suitable for arbitrary input distribution and spreading factor, which significantly improves backscatter throughput.
	This is the first paper to consider practical backscatter-primary decoding scheme to exploit backscatter modulation as dynamic passive beamforming at the co-located receiver.

@@

\begin{section}{RIScatter}
	\begin{subsection}{Principles}
		\gls{rf} wave scattering or reflecting are often manipulated by passive antennas or programmable metamaterial \cite{Liang2022}.
		The former receives the impinging signals and reradiates some back to the space, while the latter reflects at the space-cell boundary and mainly applies a phase shift.
		In the scattered signal, the structural mode component depends on the scatterer geometry and material.
		It is usually modelled as a DC offset \cite{Boyer2014} or part of environment multipath \cite{Thomas2012}.
		On the other hand, the antenna mode component depends on the impedance mismatch and is widely exploited in scattering applications.
		For an antenna (resp. metamaterial) scatterer with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
		\begin{equation}
			\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
			\label{eq:reflection_coefficient}
		\end{equation}
		where $Z_m$ is the antenna load (resp. metamaterial cell) impedance at state $m$, and $Z$ is the antenna input (resp. medium characteristic) impedance.
		Specifically,
		\begin{itemize}
			\item \gls{bc}: The scatterer is an information source with random reflection pattern over time.
			The reflection coefficient at state $m$ is used merely an information codeword \cite{Thomas2012a}
			\begin{equation}
				\Gamma_m = \alpha_m \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
				\label{eq:backscatter_modulation}
			\end{equation}
			where $\alpha_m \in \mathbb{I}$ is the amplitude scattering ratio and $c_m$ is the constellation point.
			\item \gls{ris}: The scatterer is a channel shaper with deterministic reflection pattern over time.
			The reflection coefficient at state $m$ is used merely a passive beamforming codeword \cite{Wu2018}
			\begin{equation}
				\Gamma_m = \alpha_m \exp(j \theta_m),
				\label{eq:passive_beamforming}
			\end{equation}
			where $\theta_m$ is the phase shift.\footnote{It is commonly assumed $\alpha_m \ll 1$ in \gls{bc} and $\alpha_m=1$ in \gls{ris}.}
		\end{itemize}

		RIScatter generalizes \gls{bc} and \gls{ris} from a probabilistic perspective.
		Each reflection coefficient is simultaneously an information codeword and a passive beamforming codeword.
		As shown in Fig.~\ref{fg:scatter_comparison}, RIScatter nodes semi-randomly choose reflection states according to an input distribution that strikes a balance between backscatter modulation and passive beamforming.
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Reflection Pattern]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and reflection pattern of scattering applications.
				``PB'', ``BB'', and ``CB'' refer to primary symbol block, backscatter symbol block, and channel block, respectively.
				Shadowing means presence of primary link.
				In this example, the optimal passive beamformer corresponds to state \num{2}, and the spreading factor is \num{4} for RIScatter and \num{8} for \gls{ambc}/\gls{sr}.
				\gls{bc} and \gls{ris} can be viewed as extreme cases of RIScatter, where the input distribution boils down to uniform and degenerate, respectively.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		\begin{remark}
			Unlike dynamic \gls{ris} that simply performs a time sharing between reflection states, RIScatter conveys additional information by randomizing the reflection pattern over time while still guaranteeing the probability of occurrence of each state.
			Upon successful backscatter detection, the impact of RIScatter nodes on the primary link can be modelled as dynamic passive beamforming.
		\end{remark}

@@

	% As illustrated in Fig.~\ref{fg:energy_distribution}, conditioned on different reflection state hypotheses, the accumulated energy per backscatter block follows Gamma distribution with different scale parameters \cite{Qian2017b}.
	% \begin{figure}[!t]
	% 	\centering
	% 	\resizebox{0.8\columnwidth}{!}{
	% 		\input{assets/illustration/energy_distribution.tex}
	% 	}
	% 	\caption{
	% 		\gls{pdf} of accumulated energy per backscatter block conditioned on different reflection state hypotheses.
	% 		$z$, $t$, $\mathcal{H}$ and $\mathcal{R}$ denote the accumulated energy, decision threshold, reflection state hypothesis, and decision regions, respectively.
	% 	}
	% 	\label{fg:energy_distribution}
	% \end{figure}

	% Hence, the receiver can semi-coherently (under primary uncertainty) decode RIScatter nodes from the accumulated energy, re-encode to recover the reflection patterns, and combine those with cascaded \gls{csi} to determine the multipath response and model within the primary channel.
	% Such a backscatter-primary decoding enables simultaneous backscatter modulation and dynamic passive beamforming by only one energy comparison and re-encoding during each backscatter (instead of primary) block.
	% It enjoys a three-fold benefit compared to conventional primary-backscatter decoding with \gls{sic}, namely
	% 1) reduces the re-encoding frequencies to $1/N$;
	% 2) requires no re-precoding, cancellation, or \gls{mrc};
	% 3) preserves backscatter modulation and passive beamforming at smaller spreading factors for potentially higher backscatter throughput.

	% % ! add to signal model
	% % Importantly, RIScatter system design only requires the \emph{direct} transmitter-receiver and \emph{cascaded} transmitter-scatter-receiver \gls{csi}.
	% % For dispersed RIScatter nodes, relevant \gls{csi} can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel approaches \cite{Jin2021a} originally proposed for \gls{bc}.
	% % For co-located RIScatter nodes, the estimation can be simplified by group-based \cite{Zheng2019} and hierarchical \cite{You2019} trainings originally proposed for \gls{ris}.

	\begin{subsection}{System Model}
		\label{se:system_model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.7\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{riscatter_network.pdf_tex}
			}
			\caption{A single-user multi-node RIScatter system.}
			\label{fg:riscatter_network}
		\end{figure}
		As shown in Fig.~\ref{fg:riscatter_network}, we consider a RIScatter system where a $Q$-antenna \gls{ap} serves a single-antenna user and $K$ nearby dispersed or co-located RIScatter nodes.
		All nodes have $M$ available reflection states.
		In the primary point-to-point system, the \gls{ap} transmits information to the user over a multipath channel\footnote{It is assumed the primary symbol duration is much longer than multipath delay spread (no inter-symbol interference).} enhanced by RIScatter nodes.
		In the backscatter multiple access system, the \gls{ap} acts as carrier emitter, the RIScatter nodes modulate over scattered signal, and the user jointly decodes all node messages.\footnote{It is assumed that the signal going through two or more RIScatter nodes is too weak to be received by the user.}
		For simplicity, we consider a quasi-static block fading model and focus on a specific channel block where the \gls{csi} remains constant.
		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\text{D}}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\text{B},k}$, and the cascaded \gls{ap}-node $k$-user channel as $\boldsymbol{h}_{\text{C},k}^\mathsf{H} \triangleq h_{\text{B},k} \boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$.
		We assume the direct and cascaded \gls{csi} are available at the \gls{ap} and user\footnote{The cascaded \gls{csi} can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel \cite{Jin2021a} approaches for dispersed nodes, or group-based \cite{Zheng2019} or hierarchical \cite{You2019} approaches for co-located nodes.}.
		% It is assumed that the signal going through two or more RIScatter nodes is too weak to be received by the user, the primary symbol duration is much longer than multipath delay spread (no inter-symbol interference), and the spreading factor $N$ is a positive integer.
		% For dispersed RIScatter nodes, relevant \gls{csi} can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel approaches \cite{Jin2021a} originally proposed for \gls{bc}. For co-located RIScatter nodes, the estimation can be simplified by group-based \cite{Zheng2019} and hierarchical \cite{You2019} trainings originally proposed for \gls{ris}.}.

		Let $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ be the coded backscatter symbol of node $k$ and $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all nodes.
		Due to double modulation, the composite channel is a function of backscatter symbol tuple\footnote{\eqref{eq:equivalent_channel_bc} and \eqref{eq:equivalent_channel_ris} are often used in \gls{bc} and \gls{ris} literature, respectively.}
		\begin{subequations}
			\label{eq:equivalent_channel}
			\begin{align}
				\boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \label{eq:equivalent_channel_bc}                    \\
				 & = \boldsymbol{h}_{\text{D}}^\mathsf{H} + \boldsymbol{x}^\mathsf{H} \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\text{C}}, \label{eq:equivalent_channel_ris}
			\end{align}
		\end{subequations}
		where $\alpha_k \in \mathbb{I}$ is the common amplitude scattering ratio of node $k$, $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^\mathsf{T} \in \mathbb{I}^{K}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^\mathsf{H} \in \mathcal{X}^{K}$, and $\boldsymbol{H}_{\text{C}} \triangleq [\boldsymbol{h}_{\text{C},1},\ldots,\boldsymbol{h}_{\text{C},K}]^\mathsf{H} \in \mathbb{C}^{K \times Q}$.

		Next, we consider the signal model within one backscatter block.
		The signal received by the user at primary block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is
			y[n] = \boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}}) \boldsymbol{w} s[n] + v[n],
			\label{eq:receive_signal}
		where $\boldsymbol{w} \in \mathbb{C}^{Q}$ is the active beamformer satisfying average transmit power constraint $\lVert \boldsymbol{w} \rVert^2 \le P$, $s \sim \mathcal{CN}(0,1)$ is the primary symbol, and $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn} with average power $\sigma_v^2$.
		Let $m_k \in \mathcal{M} \triangleq \{1,\ldots,M\}$ be the reflection state index of node $k$ and $m_{\mathcal{K}} \triangleq (m_1,\ldots,m_K)$ be the state index tuple of all nodes.
		% The backscatter symbol $x_k$ (resp. symbol tuple $x_{\mathcal{K}}$) is a random variable (resp. variable tuple) that takes value $x_{m_k}$ (resp. value tuple $x_{m_{\mathcal{K}}}$) when state $m_k$ (resp. state tuple $m_{\mathcal{K}}$) is employed.
		The backscatter symbol tuple $x_{\mathcal{K}}$ is a random variable that takes value $x_{m_{\mathcal{K}}}$ when state tuple $m_{\mathcal{K}}$ is used.
		Conditioned on hypothesis $\mathcal{H}_{m_{\mathcal{K}}}$, the receive signal per primary block follows \gls{cscg} distribution $\mathcal{CN}(0,\sigma_{m_{\mathcal{K}}}^2)$, where

>>>>

% glossaries-extra
\glsdisablehyper
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communication}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communication}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{clt}{CLT}{Central Limit Theorem}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmmac}{DMMAC}{Discrete Memoryless Multiple Access Channel}
\newacronym{dcmc}{DCMC}{Discrete-input Continuous-output Memoryless Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{mbc}{MBC}{Monostatic \glsentryshort{bc}}
\newacronym{bbc}{BBC}{Bistatic \glsentryshort{bc}}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{nlos}{NLoS}{Non-Line-of-Sight}
\newacronym{zf}{ZF}{Zero-Forcing}
\newacronym{mmse}{MMSE}{Minimum Mean-Square-Error}
\newacronym{fpga}{FPGA}{Field-Programmable Gate Array}
\newacronym{ber}{BER}{Bit Error Rate}

@@

	\emph{Second,} we address the first challenge and propose a practical receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern, and coherently decodes the primary link.
	We consider backscatter energy detection and formulate a \gls{dmmac} over disjoint decision regions.
	Thanks to the double modulation and symbol-level passive precoding, the semi-coherent backscatter decoding is essentially part of channel training.
	Once the backscatter information is successfully decoded, its contribution can be modelled within composite channel as dynamic passive beamforming (rather than re-precoded and cancelled).
	% It requires only one energy comparison and re-encoding per backscatter symbol and is much efficient than conventional decoding based on \gls{sic} and \gls{mrc}.
	Such a sequential decoding transforms backscatter modulation into dynamic passive beamforming using only one re-encoding and one energy level comparison per backscatter block.
	It is also suitable for arbitrary input distribution and spreading factor, which significantly improves backscatter throughput.
	This is the first paper to consider practical backscatter-primary decoding scheme to exploit backscatter modulation as dynamic passive beamforming at the co-located receiver.

@@

\begin{section}{RIScatter}
	\begin{subsection}{Principles}
		\gls{rf} wave scattering or reflecting are often manipulated by passive antennas or programmable metamaterial \cite{Liang2022}.
		The former receives the impinging signals and reradiates some back to the space, while the latter reflects at the space-cell boundary and mainly applies a phase shift.
		In the scattered signal, the structural mode component depends on the scatterer geometry and material.
		It is usually modelled as a DC offset \cite{Boyer2014} or part of environment multipath \cite{Thomas2012}.
		On the other hand, the antenna mode component depends on the impedance mismatch and is widely exploited in scattering applications.
		For an antenna (resp. metamaterial) scatterer with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
		\begin{equation}
			\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
			\label{eq:reflection_coefficient}
		\end{equation}
		where $Z_m$ is the antenna load (resp. metamaterial cell) impedance at state $m$, and $Z$ is the antenna input (resp. medium characteristic) impedance.
		Specifically,
		\begin{itemize}
			\item \gls{bc}: The scatterer is an information source with random reflection pattern over time.
			The reflection coefficient is used merely an information codeword \cite{Thomas2012a}
			\begin{equation}
				\Gamma_m = \alpha_m \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
				\label{eq:backscatter_modulation}
			\end{equation}
			where $\alpha_m \in \mathbb{I}$ is the amplitude scattering ratio and $c_m$ is the constellation point.
			\item \gls{ris}: The scatterer is a channel shaper with deterministic reflection pattern over time.
			The reflection coefficient is used merely a passive beamforming codeword \cite{Wu2018}
			\begin{equation}
				\Gamma_m = \alpha_m \exp(j \theta_m),
				\label{eq:passive_beamforming}
			\end{equation}
			where $\theta_m$ is the phase shift.\footnote{It is commonly assumed $\alpha_m \ll 1$ in \gls{bc} and $\alpha_m=1$ in \gls{ris}.}
		\end{itemize}

		RIScatter generalizes \gls{bc} and \gls{ris} from a probabilistic perspective.
		Each reflection coefficient simultaneously acts as an information codeword and a passive beamforming codeword.
		As shown in Fig.~\ref{fg:scatter_comparison}, RIScatter nodes semi-randomly choose reflection states according to an input distribution, which strikes a balance between backscatter modulation and passive beamforming.
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Reflection Pattern]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and reflection pattern of scattering applications.
				``PB'', ``BB'', and ``CB'' refer to primary symbol block, backscatter symbol block, and channel block, respectively.
				Shadowing means presence of primary link.
				In this example, the optimal passive beamformer corresponds to state \num{2}.
				The spreading factor is \num{4} for RIScatter and \num{8} for \gls{ambc}/\gls{sr}.
				\gls{bc} and \gls{ris} can be viewed as extreme cases of RIScatter, where the input distribution boils down to uniform and degenerate, respectively.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		\begin{remark}
			Unlike dynamic \gls{ris} that simply performs a time sharing between reflection states, RIScatter conveys additional information by randomizing the reflection pattern over time while still guaranteeing the probability of occurrence of each state.
			Upon successful backscatter detection, the impact of RIScatter nodes on the primary link can be modelled as dynamic passive beamforming.
		\end{remark}

@@

	\begin{subsection}{System Model}
		\label{se:system_model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.7\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{riscatter_network.pdf_tex}
			}
			\caption{A single-user multi-node RIScatter network.}
			\label{fg:riscatter_network}
		\end{figure}
		As shown in Fig.~\ref{fg:riscatter_network}, we consider a RIScatter network where a $Q$-antenna \gls{ap} serves a single-antenna user and $K$ nearby dispersed or co-located RIScatter nodes.
		All nodes have $M$ available reflection states.
		In the primary point-to-point system, the \gls{ap} transmits information to the user over a multipath channel\footnote{It is assumed the primary symbol duration is much longer than multipath delay spread (no inter-symbol interference).} enhanced by RIScatter nodes.
		In the backscatter multiple access system, the \gls{ap} acts as carrier emitter, the RIScatter nodes modulate over scattered signal, and the user jointly decodes all node messages.\footnote{It is assumed the signal going through two or more RIScatter nodes is too weak to be received by the user.}
		For simplicity, we consider a quasi-static block fading model and focus on a specific channel block where the \gls{csi} remains constant.
		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\text{D}}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\text{B},k}$, and the cascaded \gls{ap}-node $k$-user channel as $\boldsymbol{h}_{\text{C},k}^\mathsf{H} \triangleq h_{\text{B},k} \boldsymbol{h}_{\text{F},k}^\mathsf{H} \in \mathbb{C}^{1 \times Q}$.
		We assume the direct and cascaded \gls{csi} are available at the \gls{ap} and user\footnote{The cascaded \gls{csi} can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel \cite{Jin2021a} approaches for dispersed nodes, or group-based \cite{Zheng2019} or hierarchical \cite{You2019} approaches for co-located nodes.}.

		Let $\alpha_k \in \mathbb{I}$ be the amplitude scattering ratio of node $k$, $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ be the coded backscatter symbol of node $k$, and $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all nodes.
		As a sequence of double modulation, the composite channel is a function of backscatter symbol tuple\footnote{\eqref{eq:equivalent_channel_bc} and \eqref{eq:equivalent_channel_ris} are often used in \gls{bc} and \gls{ris} literature, respectively.}
		\begin{subequations}
			\label{eq:equivalent_channel}
			\begin{align}
				\boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \label{eq:equivalent_channel_bc}                    \\
				 & = \boldsymbol{h}_{\text{D}}^\mathsf{H} + \boldsymbol{x}^\mathsf{H} \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\text{C}}, \label{eq:equivalent_channel_ris}
			\end{align}
		\end{subequations}
		where $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^\mathsf{T} \in \mathbb{I}^{K}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^\mathsf{H} \in \mathcal{X}^{K}$, and $\boldsymbol{H}_{\text{C}} \triangleq [\boldsymbol{h}_{\text{C},1},\ldots,\boldsymbol{h}_{\text{C},K}]^\mathsf{H} \in \mathbb{C}^{K \times Q}$.
		Within one backscatter block, the signal received by the user at primary block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is
		\begin{equation}
			y[n] = \boldsymbol{h}^\mathsf{H}(x_{\mathcal{K}}) \boldsymbol{w} s[n] + v[n],
			\label{eq:receive_signal}
		\end{equation}
		where $\boldsymbol{w} \in \mathbb{C}^{Q}$ is the active beamformer satisfying $\lVert \boldsymbol{w} \rVert^2 \le P$, $P$ is the average transmit power budget, $s \sim \mathcal{CN}(0,1)$ is the primary symbol, and $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn} with variance $\sigma_v^2$.

		Let $m_k \in \mathcal{M} \triangleq \{1,\ldots,M\}$ be the reflection state index of node $k$, and $m_{\mathcal{K}} \triangleq (m_1,\ldots,m_K)$ be the state index tuple of all nodes.
		The backscatter symbol $x_k$ (resp. symbol tuple $x_{\mathcal{K}}$) is a random variable that takes value $x_{m_k}$ (resp. value tuple $x_{m_{\mathcal{K}}}$) with probability $p(x_{m_k})$ (resp. $p(x_{m_{\mathcal{K}}})$).
		\begin{remark}
			Dispersed RIScatter nodes encode independently such that
			\begin{equation}
				p(x_{m_{\mathcal{K}}}) = \prod_k p(x_{m_k}).
				\label{eq:equivalent_distribution}
			\end{equation}
			When the nodes are co-located, they can jointly encode by designing $p(x_{m_{\mathcal{K}}})$ directly.
		\end{remark}

		Let $z=\sum_{n} \bigl\lvert y[n] \bigr\rvert^2$ be the receive energy per backscatter block.
		When backscatter symbol tuple $x_{m_\mathcal{K}}$ occurs, the receive signal $y$ follows \gls{cscg} distribution $\mathcal{CN}(0,\sigma_{m_{\mathcal{K}}}^2)$ with variance
		\begin{equation}
			\sigma_{m_{\mathcal{K}}}^2 = \lvert \boldsymbol{h}^\mathsf{H}(x_{m_{\mathcal{K}}}) \boldsymbol{w} \rvert^2 + \sigma_v^2,
			\label{eq:receive_variance}
		\end{equation}
		and $z$ follows Gamma distribution with conditional \gls{pdf}
		\begin{equation}
			f(z|x_{m_{\mathcal{K}}}) = \frac{z^{N-1} \exp(-z/\sigma_{m_{\mathcal{K}}}^2)}{\sigma_{m_{\mathcal{K}}}^{2N} (N-1)!}.
			\label{eq:energy_distribution}
		\end{equation}

		\begin{remark}
			We have assumed Gaussian codebook for the primary source and finite support for the backscatter nodes.
			The proposed framework is extendable to non-Gaussian primary source, and the conditional \gls{pdf} \eqref{eq:energy_distribution} can be approximated using \gls{clt} for large $N$.
		\end{remark}

		The user first jointly decodes the backscatter message of all nodes using a low-complexity energy detector.
		The energy detector formulates a \gls{dmtc} of size $M^K \times M^K$.

		\begin{remark}
			The capacity-achieving decision region design for \gls{dmtc} with non-binary inputs in arbitrary distribution remains an open issue.
			It was proved deterministic detectors can be rate-optimal, but non-convex regions (comprise non-adjacent partitions) are generally required and the optimal number of thresholds is unknown \cite{Nguyen2018,Nguyen2021}.
			Next, we restrict the energy detector to convex deterministic decision regions and consider sequential threshold design.
		\end{remark}

		Let $L = M^K$ be the number of backscatter symbol tuples.
		Sort $\{\sigma_{m_{\mathcal{K}}}^2\}$ in ascending order and denote the result sequence as $\sigma_1^2,\ldots,\sigma_L^2$.
		With sequential thresholding, the decision region of backscatter symbol tuple $l \in \mathcal{L} = \{1,\ldots,L\}$ is\footnote{$m_{\mathcal{K}}$ and $l$ are one-to-one and used interchangeably in the following context.}
		\begin{equation}
			\mathcal{R}_{l} \triangleq [t_{l-1},t_l), \quad 0 \le t_{l-1} \le t_l,
		\end{equation}
		where $t_l$ is the decision threshold between hypotheses of tuple $x_{l}$ and $x_{l+1}$.
		An example is shown in Fig.~\ref{fg:energy_distribution}.
